---
title: "Comparison with Weather"
output: html_notebook
---

Import data:
```{r}
data = read.csv("C:/Users/beckr/Documents/Uni_Work/Masters/Advanced_Project/Data/Weather/Australia_summer.csv")
```

Plotting:

Weekend??


```{r}
plot(data$WEEKEND,data$Count, xlab = 'Weekend?', ylab = 'log(ETF)')
```

Difficult to tell if the weekend is significant in effecting the number of tweets - doesn't seem have the effect we'd expect. (less tweets)


Remove zero days:
```{r}
#new_data = subset(data, ETF_new != 0)
new_data = data
```



Simple Exponential Model


```{r}
#exponential.model <- lm(log(new_data$ETF_new)~ new_data$MAX)
exponential.model <- lm(log(data$ETF_new+1)~ data$MAX)

summary(exponential.model)
```

```{r}
b = summary(exponential.model)$coefficients[1]
a = summary(exponential.model)$coefficients[2]
b_e = summary(exponential.model)$coefficients[3]
a_e = summary(exponential.model)$coefficients[4]
```

```{r}
plot(data$MAX,data$ETF_new, xlab = 'Temperature (Celcius)', ylab = 'log(ETF)')

#jpeg('inexp.jpg')

plot(data$MAX,data$ETF_new, xlab = 'Temperature (Celcius)', ylab = 'ETF')

par(new = TRUE)

curve(exp(a*x+b)-1, from=-5, to=30, axes = FALSE ,  xlab = '', ylab = '', col='blue')

par(new = TRUE)
curve(exp((a-a_e)*x+b-b_e-1), from=-5, to=30, axes = FALSE , xlab = '', ylab = '',  lty=2, col='red')
par(new = TRUE)
curve(exp((a+a_e)*x+b+b_e-1), from=-5, to=30, axes = FALSE , xlab = '', ylab = '', lty=2, col = 'red')
```


General additive model:

```{r}
gam1 <- gam(ETF_new ~ s(MAX, bs="cr"), data=data)

summary(gam1)
```

```{r}
jpeg('ingam.jpg')

plot(gam1, ylab = "ETF")
```

```{r}
gam.check(gam1)
```

```{r}
plot(new_data$MAX,new_data$ETF_new, xlab = 'Temperature (Celcius)', ylab = 'log(ETF)')

#jpeg('ukexp.jpg')

plot(new_data$MAX,new_data$ETF_new, xlab = 'Temperature (Celcius)', ylab = 'ETF')

par(new = TRUE)

curve(exp(a*x+b)-1, from=-5, to=30, axes = FALSE ,  xlab = '', ylab = '', col='blue')

par(new = TRUE)
curve(exp((a-a_e)*x+b-b_e-1), from=-5, to=30, axes = FALSE , xlab = '', ylab = '',  lty=2, col='red')
par(new = TRUE)
curve(exp((a+a_e)*x+b+b_e-1), from=-5, to=30, axes = FALSE , xlab = '', ylab = '', lty=2, col = 'red')
```


Standardise:

```{r}
scaled_data = as.data.frame(scale(new_data[c(3,4,5,6,7,8,10,11,12,15)], scale = TRUE, center = FALSE))




new_data = cbind(scaled_data, new_data$WEEKEND)
```

**Initial Plots**

```{r}
pairs(new_data[c(2:11)], pch = 19, lower.panel = NULL)
```

**Models**

Multivariate Adaptive Regression Splines (MARS)


```{r}
set.seed(123)
split <- initial_split(new_data, prop = .7, strata = "ETF_new")
train <- training(split)
test  <- testing(split)
```


```{r}
mars1 <- earth(
  ETF_new ~ .,  
  data = train   
)

print(summary(mars1))
```


```{r}
# create a tuning grid
hyper_grid <- expand.grid(
  degree = 1:3, 
  nprune = floor(seq(2, 100, length.out = 10))
  )

head(hyper_grid)
```

```{r}
# for reproducibiity
set.seed(123)

# cross validated model
tuned_mars <- train(
  x = subset(train, select = -ETF_new),
  y = train$ETF_new,
  method = "earth",
  metric = "RMSE",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = hyper_grid
)

# best model
tuned_mars$bestTune
##    nprune degree
## 14     34      2

# plot results
ggplot(tuned_mars)
```

**Linear Model**

```{r}
fit <- lm(log(ETF_new+1) ~ TEMP + TEMP_3DAY + TEMP_7DAY + TEMP*DEWP + TEMP*WDSP
          + TEMP*MXSPD + MAX + MIN + TEMP*PRCP, data = new_data)

print(summary(fit))
```

```{r}

step(lm(log(ETF_new+1) ~ TEMP + TEMP_3DAY + TEMP_7DAY + TEMP*DEWP + TEMP*WDSP + TEMP*MXSPD + MAX + MIN + TEMP*PRCP, data = new_data), direction = 'both')
```

UK 
```{r}
#fit_new = lm(formula = log(ETF_new+1) ~ TEMP + #TEMP_3DAY + DEWP + MXSPD + 
#    MAX + MIN + PRCP + TEMP:DEWP + TEMP:PRCP, data = #new_data)

#Summer
fit_new = lm(formula = log(ETF_new + 1) ~ TEMP + TEMP_3DAY + DEWP + MAX + 
    MIN + PRCP + TEMP:DEWP, data = new_data)

summary(fit_new)
```

```{r}
table = tidy(fit_new)
kable(table)
```

Ireland
```{r}
#fit_new = lm(formula = log(ETF_new) ~ TEMP + WDSP + MXSPD + MAX + MIN + 
#    PRCP + TEMP:MXSPD + TEMP:PRCP, data = new_data)

fit_new =lm(formula = log(ETF_new + 1) ~ TEMP + TEMP_7DAY + DEWP + WDSP + 
    MIN + TEMP:DEWP + TEMP:WDSP, data = new_data)

summary(fit_new)
```

Australia
```{r}
#fit_new = lm(formula = log(ETF_new) ~ TEMP + DEWP + #PRCP + TEMP:DEWP + 
#    TEMP:PRCP, data = new_data)

fit_new = lm(formula = log(ETF_new + 1) ~ TEMP + TEMP_7DAY + MAX + PRCP + 
    TEMP:PRCP, data = new_data)

summary(fit_new)
```

India
```{r}
#fit_new = lm(formula = log(ETF_new) ~ TEMP + WDSP + #MAX + MIN + PRCP + 
#    TEMP:WDSP + TEMP:PRCP, data = new_data)

fit_new = lm(formula = log(ETF_new + 1) ~ TEMP + TEMP_7DAY + DEWP + WDSP + 
    MXSPD + TEMP:DEWP + TEMP:WDSP, data = new_data)

summary(fit_new)
```


```{r}
plot(fit_new)
```


Interaction terms:
We set values for all terms. We set a high value and a low value for the interaction variable with temp. We allow temp to vary and see how ETF is affected:

```{r}
temp = mean(new_data$TEMP)
temp3 = mean(new_data$TEMP_3DAY)
temp7 = mean(new_data$TEMP_7DAY)
dewp = mean(new_data$DEWP)
wdsp = mean(new_data$WDSP)
mxspd = mean(new_data$MXSPD)
max = mean(new_data$MAX)
min =mean(new_data$MIN)
prcp = mean(new_data$PRCP)
weekend = 0
```


```{r}
prcp_max = max(new_data$PRCP)
prcp_min = min(new_data$PRCP)

wdsp_max = max(new_data$WDSP)
wdsp_min = min(new_data$WDSP)

mxspd_max = max(new_data$MXSPD)
mxspd_min = min(new_data$MXSPD)

dewp_max = max(new_data$DEWP)
dewp_min = min(new_data$DEWP)
```

UK

```{r}
ETF_dmin = exp(-26.2725512 + 5.7808392*new_data$TEMP + 2.4745839*temp3 +
            14.9744978*dewp_min + 0.7735362*mxspd + 13.1000051*max +
            4.4893313*min + 1.9904747*prcp - 16.176963*new_data$TEMP*dewp_min -
            2.3409757*new_data$TEMP*prcp)

ETF_dmax = exp(-26.2725512 + 5.7808392*new_data$TEMP + 2.4745839*temp3 +
            14.9744978*dewp_max + 0.7735362*mxspd + 13.1000051*max +
            4.4893313*min + 1.9904747*prcp - 16.176963*new_data$TEMP*dewp_max -
            2.3409757*new_data$TEMP*prcp)

plot(new_data$TEMP, ETF_dmin)
plot(new_data$TEMP, ETF_dmax)
```

```{r}
ETF_pmin = exp(-26.2725512 + 5.7808392*new_data$TEMP + 2.4745839*temp3 +
            14.9744978*dewp + 0.7735362*mxspd + 13.1000051*max +
            4.4893313*min + 1.9904747*prcp_min - 16.176963*new_data$TEMP*dewp -
            2.3409757*new_data$TEMP*prcp_min)

ETF_pmax = exp(-26.2725512 + 5.7808392*new_data$TEMP + 2.4745839*temp3 +
            14.9744978*dewp + 0.7735362*mxspd + 13.1000051*max +
            4.4893313*min + 1.9904747*prcp_max - 16.176963*new_data$TEMP*dewp -
            2.3409757*new_data$TEMP*prcp_max)

plot(new_data$TEMP, ETF_pmin)
plot(new_data$TEMP, ETF_pmax)
```

Summer
```{r}
ETF_dmin = 3.8832656 - 4.629717*new_data$TEMP +                 0.6433902*temp3 - 6.2306876*dewp_min +               1.2842197*max + 0.5262339*min + 0.0596073            * prcp + 4.7423341*new_data$TEMP*dewp_min


ETF_dmax = 3.8832656 - 4.629717*new_data$TEMP +                 0.6433902*temp3 - 6.2306876*dewp_max +               1.2842197*max + 0.5262339*min + 0.0596073            * prcp + 4.7423341*new_data$TEMP*dewp_max

plot(new_data$TEMP, ETF_dmin)
plot(new_data$TEMP, ETF_dmax)
```





Ireland
```{r}
ETF_pmin = exp(-20.554282 + 32.790625*new_data$TEMP - 2.348689*wdsp + 13.75217*mxspd - 
        8.964237*max - 3.139045*min + 5.451739*prcp_min - 13.227612*new_data$TEMP -
        6.775392*new_data$TEMP*prcp_min)

ETF_pmax = exp(-20.554282 + 32.790625*new_data$TEMP - 2.348689*wdsp + 13.75217*mxspd - 
        8.964237*max - 3.139045*min + 5.451739*prcp_max - 
        13.227612*new_data$TEMP*mxspd - 6.775392*new_data$TEMP*prcp_max)

plot(new_data$TEMP, ETF_pmin)
plot(new_data$TEMP, ETF_pmax)

```

```{r}

ETF_mmin = exp(-20.554282 + 32.790625*new_data$TEMP - 2.348689*wdsp + 13.75217*mxspd_min - 
        8.964237*max - 3.139045*min + 5.451739*prcp - 
        13.227612*new_data$TEMP*mxspd_min - 6.775392*new_data$TEMP*prcp)

ETF_mmax = exp(-20.554282 + 32.790625*new_data$TEMP - 2.348689*wdsp + 13.75217*mxspd_max - 
        8.964237*max - 3.139045*min + 5.451739*prcp - 
        13.227612*new_data$TEMP*mxspd_max - 6.775392*new_data$TEMP*prcp)

plot(new_data$TEMP, ETF_mmin)
plot(new_data$TEMP, ETF_mmax)
```


Australia
```{r}
ETF_dmax = exp(4.9512199 - 0.0487092*new_data$TEMP - 0.1188323*dewp_max + 9.7076976*prcp +
            0.0080329 *new_data$TEMP*dewp_max - 0.8428927*new_data$TEMP*prcp)


ETF_dmin = exp(4.9512199 - 0.0487092*new_data$TEMP - 0.1188323*dewp_min + 9.7076976*prcp +
            0.0080329 *new_data$TEMP*dewp_min - 0.8428927*new_data$TEMP*prcp)


plot(new_data$TEMP, ETF_dmin)
plot(new_data$TEMP, ETF_dmax)
```
```{r}
ETF_pmax = exp(4.9512199 - 0.0487092*new_data$TEMP - 0.1188323*dewp + 9.7076976*prcp_max +
            0.0080329 *new_data$TEMP*dewp - 0.8428927*new_data$TEMP*prcp_max)


ETF_pmin = exp(4.9512199 - 0.0487092*new_data$TEMP - 0.1188323*dewp + 9.7076976*prcp_min +
            0.0080329 *new_data$TEMP*dewp - 0.8428927*new_data$TEMP*prcp_min)


plot(new_data$TEMP, ETF_pmin)
plot(new_data$TEMP, ETF_pmax)
```

Summer
```{r}
ETF_pmin = -1.6357237 + 0.216788*new_data$TEMP + 
            0.5617582*temp7 + 1.1593977*max +
            0.3482352*prcp_min - 
            0.4832481*new_data$TEMP*prcp_min

ETF_pmax = -1.6357237 + 0.216788*new_data$TEMP + 
            0.5617582*temp7 + 1.1593977*max +
            0.3482352*prcp_max - 
            0.4832481*new_data$TEMP*prcp_max


plot(new_data$TEMP, ETF_pmin)
plot(new_data$TEMP, ETF_pmax)
```




India
```{r}
ETF_dmin = - 12.8372207 + 12.7706336*new_data$TEMP + 0.9204825*temp7 + 10.9060787*dewp_min + 1.6921433*wdsp -0.16894*mxspd - 11.5627849*new_data$TEMP*dewp_min - 1.6083276 *new_data$TEMP*wdsp

ETF_dmax = - 12.8372207 + 12.7706336*new_data$TEMP + 0.9204825*temp7 + 10.9060787*dewp_max + 1.6921433*wdsp -0.16894*mxspd - 11.5627849*new_data$TEMP*dewp_max - 1.6083276 *new_data$TEMP*wdsp

plot(new_data$TEMP, ETF_dmin)
plot(new_data$TEMP, ETF_dmax)
```



```{r}
ETF_wmin = - 12.8372207 + 12.7706336*new_data$TEMP + 0.9204825*temp7 + 10.9060787*dewp + 1.6921433*wdsp_min -0.16894*mxspd - 11.5627849*new_data$TEMP*dewp - 1.6083276 *new_data$TEMP*wdsp_min

ETF_wmax = - 12.8372207 + 12.7706336*new_data$TEMP + 0.9204825*temp7 + 10.9060787*dewp + 1.6921433*wdsp_max -0.16894*mxspd - 11.5627849*new_data$TEMP*dewp - 1.6083276 *new_data$TEMP*wdsp_max

plot(new_data$TEMP, ETF_wmin)
plot(new_data$TEMP, ETF_wmax)
```



